Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software

### 2021-09-03 manuscript/manuscript-speed-accuracy.tex
Last wee polish of the manuscript and figures.

\usepackage[utf8]{inputenc} %unicode support
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
%\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{maori}
\graphicspath{{../figures/}}
\usepackage{xr}
\externaldocument{supplementary}

@@

\section*{Background}
Computational biology software is widely used and has produced some of
the most cited publications in the entire scientific corpus
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. These
highly-cited software tools include implementations of methods for sequence
alignment and homology inference \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
phylogenetic analysis \cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei}, 
biomolecular structure analysis \cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj}, and
visualization and data collection \cite{Kraulis1991-lt,Berman2000-to}. 
However, the popularity of a
software tool does not necessarily mean that it is accurate or
computationally efficient{\color{red};} instead usability, ease of installation,
operating system support or other indirect factors may play a greater role in a software
tool's popularity. Indeed, there have been several notable incidences where
convenient, yet inaccurate software has caused considerable harm
\cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
%Fergusson covid predictions?

@@

The gold-standard for determining accuracy is for researchers independent
of individual tool development to conduct benchmarking studies{\color{red};} these benchmarks can serve a useful role in
reducing the over-optimistic reporting of software accuracy
\cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
\cite{Norel2011-cq,Buchka:2021a}. Benchmarking typically involves the use a number of
positive and negative control datasets, so that predictions from different software tools can be
partitioned into true or false groups, allowing a variety of metrics to be
used to evaluate performance
\cite{Egan1975-nd,Hall2012-kg,Weber:2019}.
%Some benchmarks now use live, or
%frequently updated, results to indicate the latest developments in
%software performance
%\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
The aim of these benchmarks is to robustly identify tools that
make acceptable compromises in terms of balancing speed with
discriminating true and false predictions, and are therefore suited
for wide adoption by the community.

@@

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. 
Some accessible, quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing tool. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact,``large user base'', ``word-of-mouth'',
``wide-adoption'', ``personal recommendation'', and ``recommendation from a
close colleague'', were frequent responses to surveys of ``how do
scientists select software?''
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
impact:} {\color{red}many believe that} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/group
reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors may write better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as implying that slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
{\color{red}however}, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo,leiserson2020there}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}
%Other factors that are less quantifiable that influence whether a
%%software tool is selected include: whether the documentation is good,
%user-friendly, word-of-mouth and ``used in a similar analysis''
%\cite{Loman2015-bw}. However, citation metrics may be a
%useful proxy for the above.
%This may also explain the reason why
%some software continues to be used, in spite of poor relative
%performance \cite{Wadi2016-dj}.
%We note that none of these factors
%are directly linked to the top ranked features of successful software
%projects, as based upon literature surveys \cite{nasir2011critical}.
{\color{red}\textbf{6. Effective software versioning:} With the wide adoption of
public version-control systems like GitHub, quantifiable data on
software development time and intensity indicators, such as the number
of contributors to code, number of code changes and versions is
now available \cite{ray2014large,Dozmorov:2018,Mangul:2019}.}

In the following study, we explore factors that may be indicative of
software accuracy. This, in our opinion, should be one of the prime
reasons for selecting a software tool. We have mined the large and
freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks
of computational biology software, and manually extracted accuracy and
speed rankings for \numTools~ unique software tools. For
each tool, we have collected measures that may be predictive
of accuracy, and may be subjectively employed by the research
community as a proxy for software quality. These include relative
speed, relative age, the productivity and impact of the corresponding
authors, journal impact, number of citations {\color{red}and GitHub activity}.

@@

For each of the publications describing these tools, we have (where
possible) collected the journal's H5-index (
%\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{
Google Scholar Metrics
%}
), the maximum H-index and
corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
authors for each tool, and the number of times the publication(s)
associated with a tool has been cited using Google Scholar (data
collected over a 6 month period in late 2020). Note that citation metrics
are not static and will change over time. In addition, where possible we also extract the version number, the number
of commits, number of contributors
{\color{red} total number ``issues'', the proportion of issues that remain open, the number of pull requests, and the number of times the code was forked}
from public GitHub repositories.

We have computed the Spearman’s correlation coefficient for each
pairwise combination of the mean normalised accuracy and speed ranks,
with the year published, mean relative age (compared to software in the
same benchmarks), journal H5 metrics, the total number of citations,
the relative number of citations (compared to software in the same
benchmarks) and the maximum H- and corresponding M-indices for the corresponding
authors, version number,
{\color{red}and the GitHub statistics commits, contributors, pull requests, issues, \% open issues and forks}.
The results are presented in Figure~\ref{fig:allfactors}A\&B, and Additional file 1: Figures S5\&S6. We find
significant associations between most of the citation-based metrics
(journal H5, citations, relative citations, H-index and
M-index). There is also a negative correlation between the year of
publication, the relative age and many of the citation-based metrics.

Data on the number of updates to software tools from GitHub such as
the version number, {\color{red}and numbers of contributors, commits, forks and issues } was significantly
correlated with software accuracy (respective Spearman’s rhos =
{\color{red}0.15, 0.21, 0.22, 0.23, 0.23 and respective Benjamini \& Hochberg
  corrected P-values = $5.4\times 10^{-4}$, $1.1\times 10^{-3}$, $7.9\times 10^{-4}$, $3.4\times 10^{-4}$, $3.2\times 10^{-4}$, Additional file 1: Figures S6}). %See arrays "spearmansA" and "spearmansAP" in prettyPlot.R
The significance of these features was
further confirmed with a permutation test
(Figure~\ref{fig:allfactors}B).  These features were not correlated
with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Additional file 1: Figures S5\&S6}).
We also found that
reputation metrics such as citations, author and journal H-indices,
and the age of tools were generally \textbf{not} correlated with
either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).

@@

Our study finds little evidence to support that impact-based metrics
have any relationship with software quality, which is unfortunate, as these
are frequently cited reasons for selecting software tools
\cite{Loman2015-bw}. This implies that high citation
rates for bioinformatic software
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
reflection of other factors such as user-friendliness or the Matthew Effect
\cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.
%Do the authors have putative mechanism that may describe the the lack of correlation between citations and accuracy?
     {\color{red} Specifically, software tools published early are
       more likely to appear in high impact journals due to their
       perceived novelty and need. Yet without sustained maintenance
       these may be outperformed by subsequent tools, yet early
       publications still accrue citations from users, and all
       subsequent software publications as tools need to be compared
       in order to publish. Subsequent tools are not perceived to be
       as novel, hence appear in ``lower'' tier journals, despite
       being more reliable. Hence, the ``rich'' early publishers get
       richer in terms of citations. Indeed, citation counts are
       mainly predictive of age (Figure~\ref{fig:allfactors}A).}

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are over-represented at both
high and low levels of accuracy{\color{red}, with older tools enriched in this group} (Figure~\ref{fig:speedaccuracy} {\color{red}and Additional file 1: Figure S7}).
In addition, there is an large under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences as these tools may {\color{red}nevertheless} have further uses.

@@

We propose that the full spectrum of software tool accuracies and
speeds serves a useful purpose to the research community. Like
negative results, if honestly reported this information, illustrates
to the research community that certain approaches are not practical
research avenues
\cite{fanelli2012negative}. %Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
The current novelty-seeking practices of many publishers, editors,
reviewers and authors of software tools therefore may be depriving our
community of tools for building effective and productive workflows.
{\color{red} Indeed, the drive for novelty may be an actively harmful
  criteria for the software development community, just as it is for reliable
  and reproducible research
  \cite{brembs2019reliable}. Novelty-criteria for publication may, in
  addition, discourage continual, incremental improvements in code
  post-publication in favour of splashy new tools that are likely to
  accrue more citations. }

{\color{red} In addition we suggest that further efforts be made to
  encourage continual updates to software tools. To paraphrase some of
  the suggestions of Siepel (2019), these efforts may include more
  secure positions for developers, institutional promotion criteria
  include software maintenance, lower publication barriers for
  significant software updates, encourage further funding for software
  maintenance and improvement - not just new tools
  \cite{siepel2019challenges}. If these issues were recognised by
  research managers, funders and reviewers, then perhaps
  the future bioinformatic software tool landscape will be much improved. }

@@

\textbf{Criteria for inclusion:} We are interested in using
computational biology benchmarks that satisfy Boulesteix’s (ALB) three
criteria for a ``neutral comparison study''
\cite{Boulesteix2013-vb}. Firstly, the main focus of the article is
the comparison and \textbf{not} the introduction of a new tool
{\color{red}as these can be biased} \cite{Buchka:2021a}. Secondly, the
authors should be reasonably neutral, which means that the authors
should not generally have been involved in the development of the
tools included in the benchmark. Thirdly, the test data and evaluation
criteria should be sensible. This means that the test data should be
independent of data that tools have been trained upon, and that the
evaluation measures appropriately quantify correct and incorrect
predictions. {\color{red} In addition, we excluded benchmarks with too
  few tools $\le 3$, or those where the results were inaccessible (no
  supplementary materials or poor figures).}

\textbf{Literature mining:} We identified an initial list of 10 benchmark
articles that satisfy the ALB-criteria. These were identified based
upon previous knowledge of published articles and were supplemented
with several literature searches (e.g., [``benchmark'' AND ``cputime''] was
used to query both GoogleScholar and Pubmed
\cite{Sayers2010-vm,McEntyre2001-fl}). We used these articles to seed
a machine-learning approach for identifying further candidate articles
and to identify new search terms to include. {\color{red}This is outlined in Additional file 1: Figure S1. }

@@

Where
$\delta$
was a pseudo-count added for each word ($\delta = 10^{-5}$, by default),
$f_{bg}(word)$ and $f_{tr}(word)$ were the frequencies of a $word$ in
the background and training datasets respectively. Word frequencies
were computed by counting the number of times a word appears in the
pool of titles and abstracts, the counts were normalised by the total
number of words in each set. {\color{red}Additional file 1: Figure S2 shows exemplar word scores. }

Thirdly, we also collected a group of candidate benchmark articles by
mining Pubmed for articles that were likely to be benchmarks of
bioinformatic software, these match the terms: “((bioinformatics)
AND (algorithms OR programs OR software)) AND (accuracy OR assessment
OR benchmark OR comparison OR performance) AND (speed OR
time)”. Further terms used in this search were progressively added as
relevant enriched terms were identified in later iterations. The final
query is given in {\color{red} \textbf{Additional file 1}}.

@@

\textbf{Data extraction and processing:} for each article that met the
ALB-criteria and contained data on both the accuracy and speed from
their tests, we extracted ranks for each tool. {\color{red}Until
  published datasets are made available in consistent,
  machine-readable formats this step is necessarily a manual process
  -- ranks were extracted from a mixture of manuscript figures, tables
  and supplementary materials, each data source is documented in
 Additional file 2: Table S1. In addition, a variety of
  accuracy metrics are reported e.g. ``accuracy'', ``AUROC'',
  ``F-measure'', ``Gain'', ``MCC'', ``N50'', ``PPV'', ``precision'',
  ``RMSD'', ``sensitivity'', ``TPR'', ``tree error'', etc. Our
  analysis makes the necessarily pragmatic assumption that highly
  ranked tools on one accuracy metric will also be highly ranked on
  other accuracy metrics.}  Many articles contained multiple
benchmarks, in these cases we recorded ranks from each of these, the
provenance of which is stored with the accuracy metric and raw speed
and accuracy rank data for each tool {\color{red}(Additional file 2:
  Table S1)}. In line with rank-based statistics, the cases where
tools were tied were resolved by using a midpoint rank (e.g., if tools
ranked 3 and 4 are tied, the rank 3.5 was used)
\cite{Mann1947-re}. Each rank extraction was independently verified by
at least one other co-author to ensure both the provenance of the data
could be established and that the ranks were correct. The ranks for
each benchmark were then normalised to lie between 0 and 1 using the
formula $1-\frac{r-1}{n-1}$ where ‘$r$’ is a tool’s rank and ‘$n$’ is
the number of tools in the benchmark. For tools that were benchmarked
multiple times with multiple metrics (e.g., BWA was evaluated in 6
different articles
\cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd})
a mean normalised rank was used to summarise the accuracy and speed
performance.  Or, more formally:

@@

{\color{red} The distributions for each feature and permuted accuracy or speed ranks are shown in {\color{red} Additional file 1: Figures S3\&S4}. Scatter-plots for each pair of features is shown in {\color{red} Additional file 1: Figure S5}. Plots showing the sample sizes for each tool, and feature are shown in {\color{red} Additional file 1: Figure S8}, illustrates a power analysis to show what effect sizes we are likely to detect for our sample sizes. }

@@

\section*{Acknowledgements}
The authors acknowledge the valued contribution of invaluable
discussions with Anne-Laure Boulesteix, Shinichi Nakagawa, Suetonia
Palmer and Jason Tylianakis. Murray Cox, Raquel
Norel, Alexandros Stamatakis, Jens Stoye, Tandy Warnow, and 
{\color{red}Luis Pedro Coelho and three anonymous reviewers} provided
valuable feedback on drafts of the manuscript.

@@

{\color{red}Supplementary Figures S1-S8 and the neutral software benchmark reference list the accuracy and speed data is derived from [1-66].}

@@

{\color{red}Supplementary Tables S1-S7.}

>>>>

\usepackage[utf8]{inputenc} %unicode support
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
%\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
%\usepackage{maori}
%\graphicspath{{../figures/}}
\usepackage{xr}
%\externaldocument{supplementary}

%Māori characters:
\DeclareUnicodeCharacter{0100}{\=A}
\DeclareUnicodeCharacter{0101}{\=a}
\DeclareUnicodeCharacter{0112}{\=E}
\DeclareUnicodeCharacter{0113}{\=e}
\DeclareUnicodeCharacter{012A}{\=I}
\DeclareUnicodeCharacter{012B}{\=\i}
\DeclareUnicodeCharacter{014C}{\=O}
\DeclareUnicodeCharacter{014D}{\=o}
\DeclareUnicodeCharacter{016A}{\=U}
\DeclareUnicodeCharacter{016B}{\=u}

@@

\section*{Background}
Computational biology software is widely used and has produced some of
the most cited publications in the entire scientific corpus
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. These
highly-cited software tools include implementations of methods for sequence
alignment and homology inference \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
phylogenetic analysis \cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei}, 
biomolecular structure analysis \cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj}, and
visualization and data collection \cite{Kraulis1991-lt,Berman2000-to}. 
However, the popularity of a
software tool does not necessarily mean that it is accurate or
computationally efficient{\color{black};} instead usability, ease of installation,
operating system support or other indirect factors may play a greater role in a software
tool's popularity. Indeed, there have been several notable incidences where
convenient, yet inaccurate software has caused considerable harm
\cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
%Fergusson covid predictions?

@@

The gold-standard for determining accuracy is for researchers independent
of individual tool development to conduct benchmarking studies{\color{black};} these benchmarks can serve a useful role in
reducing the over-optimistic reporting of software accuracy
\cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
\cite{Norel2011-cq,Buchka:2021a}. Benchmarking typically involves the use a number of
positive and negative control datasets, so that predictions from different software tools can be
partitioned into true or false groups, allowing a variety of metrics to be
used to evaluate performance
\cite{Egan1975-nd,Hall2012-kg,Weber:2019}.
%Some benchmarks now use live, or
%frequently updated, results to indicate the latest developments in
%software performance
%\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
The aim of these benchmarks is to robustly identify tools that
make acceptable compromises in terms of balancing speed with
discriminating true and false predictions, and are therefore suited
for wide adoption by the community.

@@

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. 
Some accessible, quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing tool. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact,``large user base'', ``word-of-mouth'',
``wide-adoption'', ``personal recommendation'', and ``recommendation from a
close colleague'', were frequent responses to surveys of ``how do
scientists select software?''
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
impact:} {\color{black}many believe that} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/group
reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors may write better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as implying that slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
{\color{black}however}, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo,leiserson2020there}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}
%Other factors that are less quantifiable that influence whether a
%%software tool is selected include: whether the documentation is good,
%user-friendly, word-of-mouth and ``used in a similar analysis''
%\cite{Loman2015-bw}. However, citation metrics may be a
%useful proxy for the above.
%This may also explain the reason why
%some software continues to be used, in spite of poor relative
%performance \cite{Wadi2016-dj}.
%We note that none of these factors
%are directly linked to the top ranked features of successful software
%projects, as based upon literature surveys \cite{nasir2011critical}.
{\color{black}\textbf{6. Effective software versioning:} With the wide adoption of
public version-control systems like GitHub, quantifiable data on
software development time and intensity indicators, such as the number
of contributors to code, number of code changes and versions is
now available \cite{ray2014large,Dozmorov:2018,Mangul:2019}.}

In the following study, we explore factors that may be indicative of
software accuracy. This, in our opinion, should be one of the prime
reasons for selecting a software tool. We have mined the large and
freely accessible PubMed database \cite{Sayers2010-vm} for benchmarks
of computational biology software, and manually extracted accuracy and
speed rankings for \numTools~ unique software tools. For
each tool, we have collected measures that may be predictive
of accuracy, and may be subjectively employed by the research
community as a proxy for software quality. These include relative
speed, relative age, the productivity and impact of the corresponding
authors, journal impact, number of citations {\color{black}and GitHub activity}.

@@

For each of the publications describing these tools, we have (where
possible) collected the journal's H5-index (
%\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{
Google Scholar Metrics
%}
), the maximum H-index and
corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
authors for each tool, and the number of times the publication(s)
associated with a tool has been cited using Google Scholar (data
collected over a 6 month period in late 2020). Note that citation metrics
are not static and will change over time. In addition, where possible we also extract the version number, the number
of commits, number of contributors
{\color{black} total number ``issues'', the proportion of issues that remain open, the number of pull requests, and the number of times the code was forked}
from public GitHub repositories.

We have computed the Spearman’s correlation coefficient for each
pairwise combination of the mean normalised accuracy and speed ranks,
with the year published, mean relative age (compared to software in the
same benchmarks), journal H5 metrics, the total number of citations,
the relative number of citations (compared to software in the same
benchmarks) and the maximum H- and corresponding M-indices for the corresponding
authors, version number,
{\color{black}and the GitHub statistics commits, contributors, pull requests, issues, \% open issues and forks}.
The results are presented in Figure~\ref{fig:allfactors}A\&B, and Additional file 1: Figures S5\&S6. We find
significant associations between most of the citation-based metrics
(journal H5, citations, relative citations, H-index and
M-index). There is also a negative correlation between the year of
publication, the relative age and many of the citation-based metrics.

Data on the number of updates to software tools from GitHub such as
the version number, {\color{black}and numbers of contributors, commits, forks and issues } was significantly
correlated with software accuracy (respective Spearman’s rhos =
{\color{black}0.15, 0.21, 0.22, 0.23, 0.23 and respective Benjamini \& Hochberg
  corrected P-values = $5.4\times 10^{-4}$, $1.1\times 10^{-3}$, $7.9\times 10^{-4}$, $3.4\times 10^{-4}$, $3.2\times 10^{-4}$, Additional file 1: Figures S6}). %See arrays "spearmansA" and "spearmansAP" in prettyPlot.R
The significance of these features was
further confirmed with a permutation test
(Figure~\ref{fig:allfactors}B).  These features were not correlated
with speed however (see Figure~\ref{fig:allfactors}A {\color{black}\& Additional file 1: Figures S5\&S6}).
We also found that
reputation metrics such as citations, author and journal H-indices,
and the age of tools were generally \textbf{not} correlated with
either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).

@@

Our study finds little evidence to support that impact-based metrics
have any relationship with software quality, which is unfortunate, as these
are frequently cited reasons for selecting software tools
\cite{Loman2015-bw}. This implies that high citation
rates for bioinformatic software
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
reflection of other factors such as user-friendliness or the Matthew Effect
\cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.
%Do the authors have putative mechanism that may describe the the lack of correlation between citations and accuracy?
     {\color{black} Specifically, software tools published early are
       more likely to appear in high impact journals due to their
       perceived novelty and need. Yet without sustained maintenance
       these may be outperformed by subsequent tools, yet early
       publications still accrue citations from users, and all
       subsequent software publications as tools need to be compared
       in order to publish. Subsequent tools are not perceived to be
       as novel, hence appear in ``lower'' tier journals, despite
       being more reliable. Hence, the ``rich'' early publishers get
       richer in terms of citations. Indeed, citation counts are
       mainly predictive of age (Figure~\ref{fig:allfactors}A).}

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are over-represented at both
high and low levels of accuracy{\color{black}, with older tools enriched in this group} (Figure~\ref{fig:speedaccuracy} {\color{black}and Additional file 1: Figure S7}).
In addition, there is an large under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences as these tools may {\color{black}nevertheless} have further uses.

@@

We propose that the full spectrum of software tool accuracies and
speeds serves a useful purpose to the research community. Like
negative results, if honestly reported this information, illustrates
to the research community that certain approaches are not practical
research avenues
\cite{fanelli2012negative}. %Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
The current novelty-seeking practices of many publishers, editors,
reviewers and authors of software tools therefore may be depriving our
community of tools for building effective and productive workflows.
{\color{black} Indeed, the drive for novelty may be an actively harmful
  criteria for the software development community, just as it is for reliable
  and reproducible research
  \cite{brembs2019reliable}. Novelty-criteria for publication may, in
  addition, discourage continual, incremental improvements in code
  post-publication in favour of splashy new tools that are likely to
  accrue more citations. }

{\color{black} In addition we suggest that further efforts be made to
  encourage continual updates to software tools. To paraphrase some of
  the suggestions of Siepel (2019), these efforts may include more
  secure positions for developers, institutional promotion criteria
  include software maintenance, lower publication barriers for
  significant software updates, encourage further funding for software
  maintenance and improvement - not just new tools
  \cite{siepel2019challenges}. If these issues were recognised by
  research managers, funders and reviewers, then perhaps
  the future bioinformatic software tool landscape will be much improved. }

@@

\textbf{Criteria for inclusion:} We are interested in using
computational biology benchmarks that satisfy Boulesteix’s (ALB) three
criteria for a ``neutral comparison study''
\cite{Boulesteix2013-vb}. Firstly, the main focus of the article is
the comparison and \textbf{not} the introduction of a new tool
{\color{black}as these can be biased} \cite{Buchka:2021a}. Secondly, the
authors should be reasonably neutral, which means that the authors
should not generally have been involved in the development of the
tools included in the benchmark. Thirdly, the test data and evaluation
criteria should be sensible. This means that the test data should be
independent of data that tools have been trained upon, and that the
evaluation measures appropriately quantify correct and incorrect
predictions. {\color{black} In addition, we excluded benchmarks with too
  few tools $\le 3$, or those where the results were inaccessible (no
  supplementary materials or poor figures).}

\textbf{Literature mining:} We identified an initial list of 10 benchmark
articles that satisfy the ALB-criteria. These were identified based
upon previous knowledge of published articles and were supplemented
with several literature searches (e.g., [``benchmark'' AND ``cputime''] was
used to query both GoogleScholar and Pubmed
\cite{Sayers2010-vm,McEntyre2001-fl}). We used these articles to seed
a machine-learning approach for identifying further candidate articles
and to identify new search terms to include. {\color{black}This is outlined in Additional file 1: Figure S1. }

@@

Where
$\delta$
was a pseudo-count added for each word ($\delta = 10^{-5}$, by default),
$f_{bg}(word)$ and $f_{tr}(word)$ were the frequencies of a $word$ in
the background and training datasets respectively. Word frequencies
were computed by counting the number of times a word appears in the
pool of titles and abstracts, the counts were normalised by the total
number of words in each set. {\color{black}Additional file 1: Figure S2 shows exemplar word scores. }

Thirdly, we also collected a group of candidate benchmark articles by
mining Pubmed for articles that were likely to be benchmarks of
bioinformatic software, these match the terms: “((bioinformatics)
AND (algorithms OR programs OR software)) AND (accuracy OR assessment
OR benchmark OR comparison OR performance) AND (speed OR
time)”. Further terms used in this search were progressively added as
relevant enriched terms were identified in later iterations. The final
query is given in {\color{black} \textbf{Additional file 1}}.

@@

\textbf{Data extraction and processing:} for each article that met the
ALB-criteria and contained data on both the accuracy and speed from
their tests, we extracted ranks for each tool. {\color{black}Until
  published datasets are made available in consistent,
  machine-readable formats this step is necessarily a manual process
  -- ranks were extracted from a mixture of manuscript figures, tables
  and supplementary materials, each data source is documented in
 Additional file 2: Table S1. In addition, a variety of
  accuracy metrics are reported e.g. ``accuracy'', ``AUROC'',
  ``F-measure'', ``Gain'', ``MCC'', ``N50'', ``PPV'', ``precision'',
  ``RMSD'', ``sensitivity'', ``TPR'', ``tree error'', etc. Our
  analysis makes the necessarily pragmatic assumption that highly
  ranked tools on one accuracy metric will also be highly ranked on
  other accuracy metrics.}  Many articles contained multiple
benchmarks, in these cases we recorded ranks from each of these, the
provenance of which is stored with the accuracy metric and raw speed
and accuracy rank data for each tool {\color{black}(Additional file 2:
  Table S1)}. In line with rank-based statistics, the cases where
tools were tied were resolved by using a midpoint rank (e.g., if tools
ranked 3 and 4 are tied, the rank 3.5 was used)
\cite{Mann1947-re}. Each rank extraction was independently verified by
at least one other co-author to ensure both the provenance of the data
could be established and that the ranks were correct. The ranks for
each benchmark were then normalised to lie between 0 and 1 using the
formula $1-\frac{r-1}{n-1}$ where ‘$r$’ is a tool’s rank and ‘$n$’ is
the number of tools in the benchmark. For tools that were benchmarked
multiple times with multiple metrics (e.g., BWA was evaluated in 6
different articles
\cite{Bao2011-lv,Caboche2014-lj,Hatem2013-cs,Schbath2012-ob,Ruffalo2011-rl,Holtgrewe2011-fd})
a mean normalised rank was used to summarise the accuracy and speed
performance.  Or, more formally:

@@

{\color{black} The distributions for each feature and permuted accuracy or speed ranks are shown in {\color{black} Additional file 1: Figures S3\&S4}. Scatter-plots for each pair of features is shown in {\color{black} Additional file 1: Figure S5}. Plots showing the sample sizes for each tool, and feature are shown in {\color{black} Additional file 1: Figure S8}, illustrates a power analysis to show what effect sizes we are likely to detect for our sample sizes. }

@@

\section*{Acknowledgements}
The authors acknowledge the valued contribution of invaluable
discussions with Anne-Laure Boulesteix, Shinichi Nakagawa, Suetonia
Palmer and Jason Tylianakis. Murray Cox, Raquel
Norel, Alexandros Stamatakis, Jens Stoye, Tandy Warnow, and 
{\color{black}Luis Pedro Coelho and three anonymous reviewers} provided
valuable feedback on drafts of the manuscript.

@@

{\color{black}Supplementary Figures S1-S8 and the neutral software benchmark reference list the accuracy and speed data is derived from [1-66].}

@@

{\color{black}Supplementary Tables S1-S7.}
